### 2014-12-11 paper/paper.tex
Added intro paragraph and Astrophysics paper

\documentclass[fontsize=11pt]{scrartcl}
\usepackage[english]{babel}
\usepackage[utf8]{inputenc}				% special symbols
\usepackage{fancyhdr}					% headers and footers
\usepackage{verbatim}					% for comments
\usepackage{graphicx}					% images
\usepackage[font=footnotesize]{caption}	% smaller image label
\usepackage[usenames,dvipsnames,svgnames,table]{xcolor} 					% highlighting
\usepackage{hyperref}				    % for links (text or internet)
\usepackage{amsmath}	    			% maths
\usepackage{relsize}

\usepackage[hyperref=true,style=authoryear,backend=biber]{biblatex}
\usepackage{filecontents}
\usepackage{setspace}
\usepackage{fancyhdr}
\usepackage[ruled,vlined]{algorithm2e}

\usepackage{listings}

\lstset{language=C++,
    basicstyle=\footnotesize\ttfamily,
    keywordstyle=\bfseries\color{blue},%
    showstringspaces=false,
    breaklines=true,%
    morekeywords={include, printf},
	backgroundcolor=\color{lightgray},
    identifierstyle=\color{black},%
    stringstyle=\color{DarkOrchid},
    commentstyle=\color{Green},%
    numbers=left,
    numberstyle={\tiny \color{black}},
    numbersep=9pt,
    emph=[1]{for,if,else,switch,case,while,return},emphstyle=[1]\color{red},
	stepnumber=5,
	tabsize=2
}

%\linespread{2}

% \pagestyle{headings}

\addbibresource{bibliography.bib}

\title{Real Time Fluid Simulation \\ \smaller using Smoothed-Particle Hydrodynamics and OpenGL}

\subtitle{Computer Graphics CS 488}

\author{Michael Berg \\ Matthias Untergassmair}

\pagestyle{fancy}
\rhead{\thepage}

\begin{document}

\maketitle

\abstract{
	This paper explores the field of Smooth Particle Hydrodynamics (SPH), starting at its 
beginnings as a tool to simulate astrophysical phenomena and following its evolution and 
implementation as a way to simulate fluids such as water. We will then give a brief 
mathematical background and ensuing algorithm of our SPH simulation followed by a detailed 
explanation of it.
}

\clearpage

\section{Introduction}
%
\hspace{6 pt} Smoothed Particle Hydrodynamics (SPH) successfully simulates fluids by breaking up a 
fluid body into individual parts, or particles. These particles together form a particle 
system that simulates various gravitational forces. Fluid movement is simulated in the 
system by moving particles around any particle moved, simulating a rippling, wave effect.
\\
\hspace*{6 pt} The ripple effect is created by first calculating which particles are surrounding 
a particle moved, and these surrounding particles are moved according to the movement 
of the first particle moved. But SPH wasn't originally intended to simulate liquid 
substances, but astrophysical phenomena.
%

\section{History}
%
something
%

\section{Mathematical Background}
%
\begin{align}
	a_i^n = \frac{F_i^n}{m_i} = \dots
	\label{math:acceleration}
\end{align}

\section{The Algorithm}
%
In the following, we denote the position for the particle $i$ at time $t$ as $x_i^t$, its velocity as $v_i^t$ and its acceleration as $a_i^t$.
We omit the vector notation ($\boldsymbol{x},\boldsymbol{v},\boldsymbol{a}$) for these quantities, since the following equations are valid for the vectors as well as for each component individually.
\par\medskip
%
As suggested in \cite{generalizedwallboundary}, we use the \emph{Velocity-Verlet} time stepping scheme as follows:
%
\begin{algorithm}
	\caption{Single Timestep with Velocity Verlet Algorithm}
	\KwData{$x_i^t$, $v_i^{t - \frac{\Delta t}{2}}$, $a_i^t$, $\Delta t$}
	\KwResult{$x_i^{t + \Delta t}$, $v_i^{t + \frac{\Delta t}{2}}$, $a_i^{t + \Delta t}$}	
	$v_i^{t+\frac{\Delta t}{2}} = v_i^{t - \frac{\Delta t}{2}} + \Delta t a_i^t$ \;
	$x_i^{t+\Delta t} = x_i^t + \Delta t v_i^{t+\frac{\Delta t}{2}}$ \;	
	$a_i^{t + \Delta t} = a_i^{t + \Delta t} (x_i^{t + \Delta t}, m_i)$ from equation \ref{math:acceleration} \;
\end{algorithm}

\clearpage

\section*{Appendix: Code}

\lstinputlisting[caption=sphModel.hpp]{../sphModel.hpp}
\lstinputlisting[caption=sphModel.cpp]{../sphModel.cpp}
\lstinputlisting[caption=Simulation.cpp]{../Simulation.cpp}
\lstinputlisting[caption=simulation.vsh]{../resources/simulation.vsh}
\lstinputlisting[caption=simulation.fsh]{../resources/simulation.fsh}

\clearpage
\nocite{*}
\printbibliography

\end{document}